The future age structure of scientific disciplines

### 2025-04-15 age_structure_forecasts.qmd
Now compiling
Added more figures and edited some existing figures

# Set default colors
options(
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)
```

@@

```{r}
#| label: fig-afterdisaggregation
#| eval: true
#| fig-cap: Retirement intentions across key industries after smoothed disaggregation.
tar_read(fig5)
```

@@

```{r}
#| label: fig-future-physics
#| eval: false
#| fig-height: 3
#| fig-cap: Two example simulations from a set of 500, using historical data from 2006-2021 to forecast age-specific population distributions for 2022-2041.
p <- physics_and_astronomy |>
  ggplot() +
  aes(x = Age, y = Working, group = Year) +
  geom_line(colour = "gray")
p1 <- p +
  geom_line(data = future_physics_and_astronomy |> filter(.rep == "1"), aes(colour = Year)) +
  labs(y = "Number of working scientists") +
  scale_color_gradientn(colors = rainbow(10)[1:8]) +
  labs(
    title = "Simulation 1",
    subtitle = "Forecasted Years: 2022 - 2041"
  ) +
  scale_x_continuous(breaks = seq(20, 100, by = 10)) +
  theme_classic()  +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "none"  # Remove legend
  )

p2 <- p +
  geom_line(data = future_physics_and_astronomy |> filter(.rep == "2"), aes(colour = Year)) +
  labs(y = NULL) +  # Remove y-axis label
  scale_color_gradientn(colors = rainbow(10)[1:8]) +
  labs(
    title = "Simulation 2",
    subtitle = "Forecasted Years: 2022 - 2041"
  ) +
  scale_x_continuous(breaks = seq(20, 100, by = 10)) +
  theme_classic()  +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )

p1 + p2
```

@@

```{r}
#| label: fig-future-physics-2035
#| eval: false
#| fig-cap: Forecasted age distribution of the working population for Physics and Astronomy in 2025 and 2035, comparing changes over 10 years. Shaded areas represent 90% prediction intervals.

# 2025 and 2035
eg_forecast <- future_physics_and_astronomy |>
  filter(Year %in% c(2025, 2035)) |>
  group_by(Age) |>
  summarise(
    mean = mean(Working),
    lo = quantile(Working, prob = 0.05),
    hi = quantile(Working, prob = 0.95),
    .groups = "drop"
  )

# Overlay
ggplot(eg_forecast) +
    aes(x = Age, color = factor(Year)) +
    geom_ribbon(aes(ymin = lo, ymax = hi, fill = factor(Year)), alpha = 0.1) +
    geom_line(aes(y = mean), size = 0.75) +
    labs(
      y = "Number of working scientists",
      title = "Forecast of Working Population",
      subtitle = "Physics and Astronomy",
      color = "Forecast Year",
      fill = "Forecast Year"
    ) +
    scale_color_manual(values = c("2025" = "#66666644", "2035" = "#c14b14")) +
    scale_fill_manual(values = c("2025" = "#66666644", "2035" = "#c14b1444")) +
    scale_x_continuous(breaks = seq(20, 100, by = 10)) +
    theme_classic() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "top"
    )
```

```{r}
#| label: fig-total-future-physics
#| eval: false
#| fig-cap: Forecasted total number of working scientists for Physics and Astronomy (2022–2041), indicating the rate of change. Shaded areas represent 90% prediction intervals.

sum_physics_and_astronomy <- physics_and_astronomy |>
  as_tibble() |>
  group_by(Year) |>
  summarise(Working = sum(Working), .groups = "drop")
sum_future_physics_and_astronomy <- future_physics_and_astronomy |>
  as_tibble() |>
  group_by(Year, .rep) |>
  summarise(Working = sum(Working), .groups = "drop")  |>
  group_by(Year) |>
  summarise(
    mean = mean(Working),
    lo = quantile(Working, prob = 0.05),
    hi = quantile(Working, prob = 0.95),
    .groups = "drop"
  )
ggplot(sum_future_physics_and_astronomy) +
    aes(x = Year) +
    geom_ribbon(aes(ymin = lo, ymax = hi), fill = "#c14b1444") +
    geom_line(aes(y = mean), color = "#c14b14") +
    geom_line(data = sum_physics_and_astronomy, aes(y = Working)) +
 labs(y = "Total number of working scientists") +
  labs(
    title = "Forecast of Total Working Population",
    subtitle = "Physics and Astronomy\nForecasted Years: 2022 - 2041"
  ) +
  scale_x_continuous(breaks = seq(2010, 2040, by = 5)) +
  scale_y_continuous(labels = scales::comma_format()) +
  theme_classic()  +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

@@

```{r}
#| label: tbl-disciplines
#| tbl-cap: Classification of scientific disciplines, based on the ASCED Narrow Fields of Education within the Broad Field of Natural and Physical Sciences. The table lists their corresponding Detailed Fields.

discipline_table <- tibble(
  Narrow_Discipline = c(
    "Physics and Astronomy", "Mathematical Sciences", "Chemical Sciences",
    "Earth Sciences", "Biological Sciences", "Other Natural and Physical Sciences",
    "Natural and Physical Sciences n.f.d."
  ),
  Detailed_Field = c(
    "Physics, Astronomy.",
    "Mathematics, Statistics, Mathematical Sciences, n.e.c.",
    "Organic Chemistry, Inorganic Chemistry, Chemical Sciences, n.e.c.",
    "Atmospheric Sciences, Geology, Geophysics, Geochemistry, Soil Science, Hydrology, Oceanography, Earth Sciences, n.e.c.",
    "Biochemistry and Cell Biology, Botany, Ecology and Evolution, Marine Science, Genetics, Microbiology, Human Biology, Zoology, Biological Sciences, n.e.c.",
    "Medical Science, Forensic Science, Food Science and Biotechnology, Pharmacology, Laboratory Technology, Natural and Physical Sciences, n.e.c.",
    " "
  )
)

discipline_table |>
  kable(col.names = c("Narrow Fields", "Detailed Fields")) |>
  add_header_above(c("Broad Field: Natural and Physical Sciences" = 2), bold = TRUE) |>
  kable_styling(latex_options = c("striped"), full_width = FALSE) |>
  column_spec(2, width = "20em")
```

@@

```{r}
#| label: fig-working-population
#| eval: false
#| fig-cap: "Age distribution of the working population across scientific disciplines based on Census data from 2006, 2011, 2016, and 2021. Each panel represents a different discipline, with lines indicating the number of active scientists in the Australian labour market by age for each Census year."
#| fig-height: 5.5
disciplines_combined <- bind_rows(
  as_tibble(physics_and_astronomy) |> mutate(Discipline = "Physics and Astronomy"),
  as_tibble(mathematical_sciences) |> mutate(Discipline = "Mathematical Sciences"),
  as_tibble(chemical_sciences) |> mutate(Discipline = "Chemical Sciences"),
  as_tibble(earth_sciences) |> mutate(Discipline = "Earth Sciences"),
  as_tibble(biological_sciences) |> mutate(Discipline = "Biological Sciences"),
  as_tibble(other_natural_and_physical_sciences) |> mutate(Discipline = "Other Natural and Physical Sciences"),
  as_tibble(natural_and_physical_sciences__nfd) |> mutate(Discipline = "Natural and Physical Sciences (n.f.d.)")
)

disciplines_combined$Discipline <- factor(
  disciplines_combined$Discipline,
  levels = c("Physics and Astronomy", "Mathematical Sciences", "Chemical Sciences", "Earth Sciences", "Biological Sciences", "Other Natural and Physical Sciences", "Natural and Physical Sciences (n.f.d.)")
)

disciplines_combined |>
  filter(Year %in% years) |>
  mutate(Year = factor(Year, levels = rev(years))) |>
  ggplot() +
  aes(x = Age, y = Working, color = Year, group = Year) +
  geom_line() +
  scale_color_manual(values = cols, name = "Census Year") +
  labs(
    y = "Number of active scientists",
    title = "Working Population by Discipline",
    subtitle = "Natural and Physical Sciences"
  ) +
  scale_x_continuous(breaks = seq(20, 100, by = 10)) +
  facet_wrap(~Discipline, scales = "free_y") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "top"
  )
```

@@

```{r}
#| label: fig-interpolated-working-population
#| fig-cap: "Age distribution of the working population across scientific disciplines, interpolated from Census data between 2006 and 2021. Each panel represents a different discipline, with lines indicating the number of active scientists in the Australian labour market by age and year."
#| fig-height: 6
#| eval: false
ggplot(disciplines_combined) +
  aes(x = Age, y = Working, group = Year, color = Year) +
  geom_line() +
  facet_wrap(~Discipline, scales = "free_y") +
  scale_x_continuous(breaks = seq(20, 100, by = 10)) +
  scale_color_gradientn(
    colours = rainbow(10)
  ) +
  labs(
    y = "Number of active scientists",
    title = "Interpolated Working Population by Discipline",
    subtitle = "Natural and Physical Sciences\n2006 - 2021"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "top"
  )
```

@@

```{r}
#| label: fig-graduates-disciplines
#| eval: false
#| fig-cap: "Number of graduates across scientific disciplines from 2006 to 2023. Each panel represents a different discipline, showing the trend in graduate numbers over time."

course_leavers_arranged <- course_leavers_lvl4_science
course_leavers_arranged$discipline <- course_leavers_lvl4_science$discipline |>
  fct_recode(
    "Natural and Physical Sciences (n.f.d.)" = "Natural and Physical Sciences, nfd"
  ) |>
  fct_relevel(
    "Physics and Astronomy",
    "Mathematical Sciences",
    "Chemical Sciences",
    "Earth Sciences",
    "Biological Sciences",
    "Other Natural and Physical Sciences",
    "Natural and Physical Sciences (n.f.d.)"
  )

ggplot(course_leavers_arranged) +
  aes(x = Year, y = Graduates) +
  geom_line() +
  facet_wrap(~discipline, scales = "free_y") +
  labs(
    y = "Number of graduates",
    title = "Graduates by Discipline",
    subtitle = "Natural and Physical Sciences\n2006 - 2023"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "none"
  )
```

@@

```{r}
#| label: fig-net-migration-disciplines
#| eval: false
#| fig-cap: Estimated net migration by age across scientific disciplines (2006–2020). Each panel represents a different discipline, showing net migration trends across age groups.
#| fig-height: 6
ggplot(disciplines_combined) +
  aes(x = Age, y = Migrants, group = Year, color = Year) +
  geom_line() +
  facet_wrap(~Discipline, scales = "free_y") +
  scale_x_continuous(breaks = seq(10, 100, by = 10)) +
  scale_color_gradientn(
    colours = rainbow(10)
  ) +
  labs(
    y = "Number of net migrants",
    title = "Estimated Net Migration by Discipline",
    subtitle = "Natural and Physical Sciences\n2006 - 2020"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "top"
  )
```

@@

```{r}
#| label: fig-forecast-working-population
#| eval: false
#| fig-cap: "Projected age distribution of the working population in 2025 and 2035 across scientific disciplines. The shaded regions represent the 90% prediction intervals, and the solid lines indicate the mean estimates for each year."
#| fig-height: 5.5
future_disciplines_combined <- bind_rows(
  as_tibble(future_physics_and_astronomy) |> mutate(Discipline = "Physics and Astronomy"),
  as_tibble(future_mathematical_sciences) |> mutate(Discipline = "Mathematical Sciences"),
  as_tibble(future_chemical_sciences) |> mutate(Discipline = "Chemical Sciences"),
  as_tibble(future_earth_sciences) |> mutate(Discipline = "Earth Sciences"),
  as_tibble(future_biological_sciences) |> mutate(Discipline = "Biological Sciences"),
  as_tibble(future_other_natural_and_physical_sciences) |> mutate(Discipline = "Other Natural and Physical Sciences"),
  as_tibble(future_natural_and_physical_sciences__nfd) |> mutate(Discipline = "Natural and Physical Sciences (n.f.d.)")
)

future_disciplines_combined$Discipline <- factor(
  future_disciplines_combined$Discipline,
  levels = c("Physics and Astronomy", "Mathematical Sciences", "Chemical Sciences", "Earth Sciences", "Biological Sciences", "Other Natural and Physical Sciences", "Natural and Physical Sciences (n.f.d.)")
)

# 2025 and 2035
eg_forecast <- future_disciplines_combined |>
  filter(Year %in% c(2025, 2035)) |>
  group_by(Discipline, Age, Year) |>
  summarise(
    mean = mean(Working),
    lo = quantile(Working, prob = 0.05),
    hi = quantile(Working, prob = 0.95),
    .groups = "drop"
  )

# Overlay
ggplot(eg_forecast) +
    aes(x = Age, color = factor(Year)) +
    geom_ribbon(aes(ymin = lo, ymax = hi, fill = factor(Year)), alpha = 0.1) +
    geom_line(aes(y = mean), size = 0.75) +
    labs(
      y = "Number of working scientists",
      title = "Forecast of Working Population by Discipline",
      subtitle = "Natural and Physical Sciences",
      color = "Forecast Year",
      fill = "Forecast Year"
    ) +
    scale_color_manual(values = c("2025" = "#66666644", "2035" = "#c14b14")) +
    scale_fill_manual(values = c("2025" = "#66666644", "2035" = "#c14b1444")) +
    scale_x_continuous(breaks = seq(20, 100, by = 10)) +
    theme_classic() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "top"
    ) +
    facet_wrap(~Discipline, scales = "free_y")
```

@@

-   **Physics and Astronomy** *(Aging Workforce):* The 35–45 age group shows a higher average than the previous 2025 cohort. There is a growing number of scientists aged 45 to 60, and more individuals aged 60+ are remaining in the workforce longer. These patterns indicate a gradual transition toward an older age structure, reflecting an aging workforce.
-   **Mathematical Sciences** *(Expansive Workforce):* Workforce growth is evident across a broad range of ages, particularly among individuals aged 25 to 70. The most significant increase occurs before 60, though there is also a clear rise in older workers (60+) remaining in the workforce rather than retiring. Overall, workforce growth is strong and evident across age groups.
-   **Chemical Sciences** *(Stationary Workforce):* The average workforce size remains stable, with a relatively narrow prediction interval, indicating a predictable pattern of stability over time. While there is a minor increase in the number of workers aged 70, this change is not substantial enough to suggest a broader shift in workforce dynamics.
\newpage
-   **Earth Sciences** *(Aging Workforce):* Early-career workforce projections are more uncertain relative to its size. A decline in the working population is observed in the mid-thirties, followed by an increase throughout the forties, though both remain modest. While the overall workforce remains relatively stable, these patterns indicate a slow shift toward an older age distribution.
-   **Biological Sciences** *(Aging Workforce):* The number of workers aged 50 to 65 is increasing, indicating growth in later-career participation.
-   **Other Natural and Physical Sciences** *(Expansive Workforce):* Workforce growth is strong across ages 30 to 70, characteristic of an expanding workforce.
-   **Natural and Physical Sciences (n.f.d.)** *(Aging Workforce):* Growth is primarily concentrated in the 55–70 age range, indicating delayed retirement and an aging workforce.

@@

```{r}
#| label: fig-forecast-total-working-scientists
#| eval: false
#| fig-cap: "Projected total number of working scientists across scientific disciplines from 2022 to 2041. The shaded region represents the 90% prediction interval, the coloured line indicates the mean estimate, and the black line represents historical data."

sum_disciplines_combined  <- disciplines_combined |>
  group_by(Discipline, Year) |>
  summarise(Working = sum(Working), .groups = "drop")

sum_future_disciplines_combined <- future_disciplines_combined |>
  group_by(Discipline, Year, .rep) |>
  summarise(Working = sum(Working), .groups = "drop")  |>
  group_by(Discipline, Year) |>
  summarise(
    mean = mean(Working),
    lo = quantile(Working, prob = 0.05),
    hi = quantile(Working, prob = 0.95),
    .groups = "drop"
  )

ggplot(sum_future_disciplines_combined) +
  aes(x = Year) +
  geom_ribbon(aes(ymin = lo, ymax = hi), fill = "#c14b1444") +
  geom_line(aes(y = mean), color = "#c14b14") +
  geom_line(data = sum_disciplines_combined, aes(y = Working)) +
  labs(
    y = "Total number of working scientists (thousands)",
    title = "Forecast of Total Working Population by Discipline",
    subtitle = "Natural and Physical Sciences\nForecasted Years: 2022 - 2041"
  ) +
  scale_x_continuous(breaks = seq(2010, 2040, by = 10)) +
  scale_y_continuous(labels = scales::label_number(scale = 1e-3)) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  facet_wrap(~Discipline, scales = "free_y")
```

>>>>

# Set default colors
options(
  ggplot2.discrete.colour = c(
    "#D55E00",
    "#0072B2",
    "#009E73",
    "#CC79A7",
    "#E69F00",
    "#56B4E9",
    "#F0E442"
  ),
  ggplot2.discrete.fill = c(
    "#D55E00",
    "#0072B2",
    "#009E73",
    "#CC79A7",
    "#E69F00",
    "#56B4E9",
    "#F0E442"
  )
)
```

@@

@@

```{r}
#| label: fig-future-physics
#| eval: true
#| fig-height: 3
#| fig-cap: Two example simulations from a set of 500, using historical data from 2006-2021 to forecast age-specific population distributions for 2022-2041.
tar_read(fig16)
```

@@

```{r}
#| label: fig-future-physics-2035
#| eval: true
#| fig-cap: Forecasted age distribution of the working population for Physics and Astronomy in 2025 and 2035, comparing changes over 10 years. Shaded areas represent 90% prediction intervals.
tar_read(fig17)
```

```{r}
#| label: fig-total-future-physics
#| eval: true
#| fig-cap: Forecasted total number of working scientists for Physics and Astronomy (2022–2041), indicating the rate of change. Shaded areas represent 90% prediction intervals.
tar_read(fig18)

```

@@

```{r}
#| label: tbl-disciplines
#| eval: true
#| tbl-cap: Classification of scientific disciplines, based on the ASCED Narrow Fields of Education within the Broad Field of Natural and Physical Sciences. The table lists their corresponding Detailed Fields.
tar_read(discipline_table)
```

@@

```{r}
#| label: fig-working-population
#| eval: true
#| fig-cap: "Age distribution of the working population across scientific disciplines based on Census data from 2006, 2011, 2016, and 2021. Each panel represents a different discipline, with lines indicating the number of active scientists in the Australian labour market by age for each Census year."
#| fig-height: 5.5
tar_read(fig19)
```

@@

```{r}
#| label: fig-interpolated-working-population
#| eval: true
#| fig-cap: "Age distribution of the working population across scientific disciplines, interpolated from Census data between 2006 and 2021. Each panel represents a different discipline, with lines indicating the number of active scientists in the Australian labour market by age and year."
#| fig-height: 6
tar_read(fig20)
```

@@

```{r}
#| label: fig-graduates-disciplines
#| eval: true
#| fig-cap: "Number of graduates across scientific disciplines from 2006 to 2023. Each panel represents a different discipline, showing the trend in graduate numbers over time."
tar_read(fig21)
```

```{r}
#| label: fig-grad-forecasts
#| eval: true
#| fig-cap: "Forecast number of graduates by discipline, 2024-2041, with 90% prediction intervals, based on historical data from 2006--2023."
tar_read(fig_grad_forecasts)
```

@@

```{r}
#| label: fig-net-migration-disciplines
#| eval: true
#| fig-cap: Estimated net migration by age across scientific disciplines (2006–2020). Each panel represents a different discipline, showing net migration trends across age groups.
#| fig-height: 6
tar_read(fig22)
```

@@

```{r}
#| label: fig-forecast-working-population
#| eval: true
#| fig-cap: "Projected age distribution of the working population in 2025 and 2035 across scientific disciplines. The shaded regions represent the 90% prediction intervals, and the solid lines indicate the mean estimates for each year."
#| fig-height: 5.5
# 2025 and 2035
tar_read(fig23)
```

@@

-   **Physics and Astronomy** *(Aging Workforce):* The 35–45 age group shows a higher average than the previous 2025 cohort. There is a growing number of scientists aged 45 to 60, and more individuals aged 60+ are remaining in the workforce longer. These patterns indicate a gradual transition toward an older age structure, reflecting an aging workforce.
-   **Mathematical Sciences** *(Expansive Workforce):* Workforce growth is evident across a broad range of ages, particularly among individuals aged 25 to 70. The most significant increase occurs before 60, though there is also a clear rise in older workers (60+) remaining in the workforce rather than retiring. Overall, workforce growth is strong and evident across age groups.
-   **Chemical Sciences** *(Stationary Workforce):* The average workforce size remains stable, with a relatively narrow prediction interval, indicating a predictable pattern of stability over time. While there is a minor increase in the number of workers aged 70, this change is not substantial enough to suggest a broader shift in workforce dynamics.
-   **Earth Sciences** *(Aging Workforce):* Early-career workforce projections are more uncertain relative to its size. A decline in the working population is observed in the mid-thirties, followed by an increase throughout the forties, though both remain modest. While the overall workforce remains relatively stable, these patterns indicate a slow shift toward an older age distribution.
-   **Biological Sciences** *(Aging Workforce):* The number of workers aged 50 to 65 is increasing, indicating growth in later-career participation.
-   **Other Natural and Physical Sciences** *(Expansive Workforce):* Workforce growth is strong across ages 30 to 70, characteristic of an expanding workforce.
-   **Natural and Physical Sciences (n.f.d.)** *(Aging Workforce):* Growth is primarily concentrated in the 55–70 age range, indicating delayed retirement and an aging workforce.

@@

```{r}
#| label: fig-forecast-total-working-scientists
#| eval: true
#| fig-cap: "Projected total number of working scientists across scientific disciplines from 2022 to 2041. The shaded region represents the 90% prediction interval, the coloured line indicates the mean estimate, and the black line represents historical data."
tar_read(fig24)
```
